---
title: "Student Report (Week 07 R Markdown Demo)"
class: "R for Data Science"
author: "Sami Cemek"
output: 
  html_document:
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mosaic)
library(readr)

# Import data
file <- "CollegeScorecard2019.csv"
college <- read_csv(file)

ncf_val <- college %>%
  filter(isNCF == TRUE)
```

## Introduction

This is a brief report that explores college debt by institution and
region for the year 2019 or 2020. We will focus on:

1.  Summarizing and visualizing one numerical variable, as well as
    interpreting the results of those summaries and graphs.
2.  Comparing distributions of a numerical variable that has been split
    into groups (categories) according to a second variable.

Suppose you are intending to present this information at a meeting of
other NCF students. Some students will have statistics or R knowledge
and some won't; you want to be able to focus on the results, but still
demonstrate transparency if anyone asks about how your code works.

You also want to be able to switch between 2019 and 2020 data easily,
since people might want to see the past year or the current year,
depending on their interest.

## Data prep

The data we will be working with comes from the College Scorecard
website, which you can access here:
<https://collegescorecard.ed.gov/data/>. The data was updated in summer
2019 and contains dozens of variables measured about thousands of
colleges across the country.

We have downloaded a smaller version of the data set for this week's
report. We have focused in on a few variables, though you are welcome to
look at the larger data set if you are interested in the topic.

Even with the smaller data set, we need to do some work before we can
start making graphs of finding numerical summaries. Here is an overview
of the variables we will be working with.

The `REGION` variable records the region of US from where each college
or institution is located. It consists of

+--------------+-----------------+-------------------------------------+
| Abbreviation | Region          | States/Areas                        |
+==============+=================+=====================================+
| GL           | Great Lakes     | (IL, IN, MI, OH, WI)                |
+--------------+-----------------+-------------------------------------+
| ME           | Mid East        | (DE, DC, MD, NJ, NY, PA)            |
+--------------+-----------------+-------------------------------------+
| NE           | New England     | (CT, ME, MA, NH, RI, VT)            |
+--------------+-----------------+-------------------------------------+
| Outlying     | Outlying Area   | (AS, FM, GU, MH, MP, PR, PW, VI)    |
+--------------+-----------------+-------------------------------------+
| Plains       | Plains          | (IA, KS, MN, MO, NE, ND, SD)        |
+--------------+-----------------+-------------------------------------+
| Rockies      | Rocky Mountains | (CO, ID, MT, UT, WY)                |
+--------------+-----------------+-------------------------------------+
| SE           | Southeast       | (AL, AR, FL, GA, KY, LA, MS, NC,    |
|              |                 | SC, TN, VA, WV)                     |
+--------------+-----------------+-------------------------------------+
| SW           | Southwest       | (AZ, NM, OK, TX)                    |
+--------------+-----------------+-------------------------------------+
| U.S.S.S.     | U.S. Service    |                                     |
|              | schools         |                                     |
+--------------+-----------------+-------------------------------------+
| West         | Far West        | (AK, CA, HI, NV, OR, WA)            |
+--------------+-----------------+-------------------------------------+

In addition, we will look at three other variables as well.
DEBT_MDN_SUPP is the median debt per institution, which has been
suppressed for sample sizes less than 30. COSTT4_A is the annual total
cost of attendance. RPY_3YR_RT_SUPP is the 3rd year repayment rate,
again suppressed for small sample sizes.

## Results

The results section is broken into 4 sub-sections. First, we look at
overall median debt across all schools. Then we break the analysis of
cost down by region, take a closer look at students' abilities to repay
their debts, and finally we examine at how NCF compares to other
schools.

### Median Debt for 2019, across all schools

Let's consider the median debt at the schools in the data set. Here are
numerical summaries and graphs for the distribution of debt. Describe
the main features of the distribution.

```{r, warning=FALSE}

favstats(~DEBT_MDN_SUPP, data = college)
gf_histogram(~DEBT_MDN_SUPP, data = college, bins = 30, fill = "steelblue", alpha = 0.8) +
  labs(title = paste("Median Debt")) + theme_light()

```

### Cost by Region

Let's break the analysis of cost down by region. Here are numerical
summaries and graphs of total cost, by region. Compare the costs between
regions. (Also pay attention to sample size!). The red line marks NCF's
cost for comparison, which was [FILL IN NCF'S COST HERE].

```{r, warning=FALSE}
favstats(COSTT4_A ~ REGION, data = college)
gf_boxplot(COSTT4_A ~ REGION, data = college, fill = "steelblue")%>%
        gf_hline(color = "red", 
                yintercept = unlist(ncf_val[1, "COSTT4_A"]), 
                data = NA) + 
  theme_light() + 
  labs(title = paste("Cost by region"))


```

As we can see from the above summaries, the SouthEast (where NCF is
located) had some of the highest costing colleges, with outliers far
above the rest of the institutions. The schools with the overall highest
and lowest cost in 2019 were:

```{r, warning=FALSE}
college %>%
  filter(COSTT4_A %in% c(min(COSTT4_A, na.rm = TRUE), max(COSTT4_A, na.rm = TRUE))) %>%
  mutate(COST = paste0("$", COSTT4_A)) %>%
  select(INSTNM, REGION, COST) 
```

### Debt Repayment

Let's consider students' ability to repay their debt within 3 years,
broken down by region. The following show summaries and graphs to
describe the distribution of repayment rates by region. Compare the
regions. (Also pay attention to sample size!)

The red line marks NCF's repayment rate for comparison, which was [FILL
IN NCF'S VALUE HERE].

```{r, warning=FALSE}
favstats(RPY_3YR_RT_SUPP~REGION, data = college)
gf_boxplot(RPY_3YR_RT_SUPP~REGION, data = college, fill = "steelblue")%>%
        gf_hline(color = "red", 
                yintercept = unlist(ncf_val[1, "RPY_3YR_RT_SUPP"]), 
                data = NA) + 
  theme_light() + 
  labs(title = paste("3-year repayment by region"))

```

Here, we see a large amount of variation in the schools across all
regions. The schools with the lowest and highest repayment rates are
shown below:

```{r, warning=FALSE}
college %>%
  filter(RPY_3YR_RT_SUPP %in% c(min(RPY_3YR_RT_SUPP, na.rm = TRUE), max(RPY_3YR_RT_SUPP, na.rm = TRUE))) %>%
  mutate(REPAY_RATE = paste0(round(100*RPY_3YR_RT_SUPP,2), "%")) %>%
  select(INSTNM, REGION, REPAY_RATE)
```

### A Closer look at New College

Now look more deeply into the data. Find New College in the data set.
Look at the 6 year median earnings and median debt. How does New College
stack up to other schools across the country and in the Southeast?

```{r, echo = FALSE, warning=FALSE}

college %>%
  filter(INSTNM == "New College of Florida") %>%
  select(INSTNM, REGION, COSTT4_A, DEBT_MDN_SUPP, RPY_3YR_RT_SUPP) 

```

Let's look at the relationship between debt and cost more closely, to
see how NCF compares. The red asterisk marks New College.

```{r, warning=FALSE }
gf_point(RPY_3YR_RT_SUPP ~ COSTT4_A, data = college, color = "steelblue", alpha = 0.8) %>%
        gf_point(RPY_3YR_RT_SUPP ~ COSTT4_A, color = "red", shape = 8, data = ncf_val)+ 
  theme_light() + 
  labs(title = paste("3-year repayment vs cost"))
```

The graph shows that NCF has a lower cost and higher repayment rate than
most colleges in the U.S.

![New College of Florida All rights
reserved.](https://image.pitchbook.com/fdR7NGmFpBsonHtc5pCs98GktXn1605394343914_200x200)
